---
title: "Peat"
output: html_document
---
#'Read function
```{r setup, include=FALSE}

library(knitr)
library(SoilDataR) #library(devtools); install_github("ktoddbrown/soilDataR")
library(ggplot2) #make pretty plots
library(plyr) #data management
library(dplyr)
library(knitr)
#mapping librarys to help with global/regional plots
library(ggmap)
library(maps)
library(mapdata)
library(fiftystater)
library(hexbin)
```

```{r loadData}
#'Change hard path to relative pathway for easy sharing
source('~/soilDataR/R/processData_Treat2015.R')
meta<- processData_Treat2015(dir='/Treat_2015', verbose=FALSE)
```


# Site locations
## Lat-lon map
```{r mapSites}

mapWorld <- borders("world", colour="gray80", fill="gray80") # create a layer of borders
#ggplot() + mapWorld
ggplot(unique(meta$site[, c('lat', 'long')])) +
  mapWorld + 
  geom_hex(aes(x=long, y=lat), bins=200) + 
  scale_fill_gradient(trans='log10')
```

## Sites with country but no lat-lon
```{r countryMap}
countryCounts <- ddply(subset(meta$site, !is.na(lat+long)), 'country',  summarize, count=length(unique(site_name)))
names(countryCounts)[1] <- 'region'
countryCounts$region <- as.character(countryCounts$region)
countryCounts$region[grepl('united States', countryCounts$region)] <- 'USA'

map.world <- map_data(map="world")

cat('These countries are unknown to the map\n')
kable(subset(countryCounts, ! region %in% map.world$region))
missingCountries <- data.frame(region=setdiff(map.world$region, countryCounts$value), count=NA)
countryCounts <- rbind(countryCounts, missingCountries)
#map.world <- merge(map.world, countryCounts, all=TRUE)
ggplot(countryCounts, aes(map_id = region)) +
  geom_map(aes(fill=count), map=map.world) +
  expand_limits(x = map.world$long, y = map.world$lat) +
  coord_cartesian(xlim=c(-180,180), ylim=c(-90,90)) +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  scale_fill_gradient(trans='log10') +
  labs(x = "", y ="") +
  theme(legend.position = "bottom", 
        panel.background = element_blank())
```

`r nrow(unique(subset(meta$site, (country=='Unknown' | country == '') & is.na(lat+long), select=c('siteID'))))` sites with no known country nor lat-lon.

## Site count by country map (include sites w/ lat-lon)
```{r totCountryCount}
countryCounts <- ddply(meta$site, 'country', summarize, count=length(unique(siteID)))
names(countryCounts)[1] <- 'region'
countryCounts$region <- as.character(countryCounts$region)
countryCounts$region[grepl('United States', countryCounts$region)] <- 'USA'

map.world <- map_data(map="world")

missingCountries <- data.frame(region=setdiff(map.world$region, countryCounts$region), count=NA)
setdiff(countryCounts$region, map.world$region)
countryCounts <- rbind(countryCounts, missingCountries)
#map.world <- merge(map.world, countryCounts, all=TRUE)
ggplot(countryCounts, aes(map_id = region)) +
  geom_map(aes(fill=count), map=map.world) +
  expand_limits(x = map.world$long, y = map.world$lat) +
  coord_cartesian(xlim=c(-180,180), ylim=c(-90,90)) +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  scale_fill_gradient(trans='log10') +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", 
        panel.background = element_blank())
```

# Summary of measurements

The following variables are listed in 
```{r types}
table(meta$meta$dataset_description)
```
All measurements are on oven dry soils and in mass percent where percent noted.

## Soil organic carbon stock

+ *soc* calculated soil carbon stock of layer/profile

```{r SOC, eval=FALSE, include=FALSE}
ggplot(merge(subset(meta$sample, grepl('^soc', meta$layer$layer_name) & '^soc' > 0),
             meta$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~units)

summary(subset(meta$sample, grepl('^soc', variable) & (value < 0)))
```

## Bulk density

+ *bd_samp* is the bulk density of the <2mm (fine earth) fraction, in which the mass is expressed on an oven-dry (105 deg) basis. This is the best form of Db for the most uses.
+ *bd_tot* is the whole soil bulk density, includes fine earth fraction and rocks.
+ *bd_whole* is the bulk density of the whole soil (coarse frags, if present, included), expressed on a 1/3 bar moisture content (a little drier than typical site moist, but not oven-dry). It is collected by the clod method. For a variety of reasons, including the presence of water, possible coarse frags, and the tendency of clods to cleave along pore spaces, this will overestimate Db relative to bd_samp. The SSL standard method here is 4A1d if you want to look it up.
+ *bd_other* is, for data contributed by NRCS SSL, the bulk density of the fine earth fraction, but expressed on a site moist condition. It is collected by a volumetric core, but given the presence of some water, will probably usually be an overestimate relative to the bd_samp. This corresponds to the SSL standard method 4A3a. For datasets contributed by sources other than NRCS, bd_other is anyone's guess. Hopefully specified in the associated bd_method variable, or the metadata sheet for that dataset.

Reasonable BD minimum 0.03 or lower and maximum 2.7 g cm-3. 
```{r BD}
plot.df <- merge(subset(meta$sample, grepl('^bd_', variable) & 
                          value > 0 & value < 2.7), meta$measurement)
ggplot(plot.df) +
  geom_histogram(aes(x=value, y=..density..)) + 
  #scale_x_log10() + 
  facet_wrap(type~units)

summary(subset(meta$sample, grepl('^bd_', variable) & 
                          (value <= 0 | value > 2.7)))

nrow(plot.df)
```

Number of samples for BD `r length(plot.df$value)`.

## Carbon (mostly organic)

+ *c_tot* carbon concentration in a dry combustion analysis, sometimes include inorganic carbon
+ *oc* organic carbon concentration from either acidified (to remove carbonates) and then ran on dry combustion analyizer, or (more common) wet oxidation 'Walkly-Black' dicromate oxidation, chemical digest for organic carbon; old methods 1930-1990s tends to over estimate it in certain soils. pH below 7 c_tot generally == oc otherwise you might have carbonates
+ *loi* loss on ignition uncorrected for C:organics, generally close to half for O-horizon, C about a third of the LOI for lower horizons.

```{r C_OC_LOI}
ggplot(merge(subset(meta$sample, grepl('^(c_|oc|loi|soc)', variable) & value > 0),
             meta$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~units)

summary(subset(meta$sample, grepl('^(c_|oc|loi)', variable) & (value < 0)))
```


## Nitrogen
+ *n_tot* most N is organic mass percent dry weight, probably estimated by dry combustion (Dumas methods), sometimes also Kjeldahl digestion converts N to nitrate and then measures nitrate.

```{r N}
ggplot(merge(subset(meta$sample, grepl('^n_', variable) & value > 0),
             meta$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~units)
```